---
title: "Covid91 vaccine study FinalFall2022"
author: "Wilhelmus Elersich"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$ The vaccine covid921 is ineffective at preventing the contraction of Covid-19 in males.

$H_1$ The vaccine covid921 is effective at preventing the contraction of Covid-19 in males.

### Wrangling Data
```{r}
Males<-filter(FinalData,sex== "M") %>%
  select("treatment","infected")
```

Here I filtered the dataset so that only males, our subpopulation of interest, remain. I also selected only the columns needed to determine the effecicacy of the vaccine, resulting in a tidier dataset than I otherwise would have.

### Graphical Results
Since we are trying to explain a categorical variable, "infected", by another categorical variable, "treatment", we should use a boxplot to visualize the relationship between the two variables. we are also more interested in percentages than quantities.

```{r}
barchartGC(~infected + treatment,data=Males, type="percent")
```
Here we can see that the percentage of those who were vaccinated contracted covid-19 is significantly smaller than those who had taken placebo. Of those who did not contract covid-19, roughly half were vaccinated, while half were not.

### Numerical Results
We shall create a table showing the data, in a similar manner to the above plot.
```{r}
tableM <- xtabs(~infected + treatment,data=Males)
rowPerc(tableM)
```
Here we can see the same results as the bar plot, but more precisely. Of those who were infected with covid-19, 68.44% had taken placebo, while 31.56 had taken the vaccine. Of those who remained uninfected, close to half (49.88%) had taken placebo, while the other "half" (50.12%) had taken the vaccine.

### Inferential Results


## Females



## LGBTQ



## Druggies


# Overall Results and Conclusions