---
title: "Covid91 vaccine study FinalFall2022"
author: "Wilhelmus Elersich"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$ The vaccine covid921 is has no effect on the contraction of Covid-19 in males.

$H_1$ The vaccine covid921 is has an effect on the contraction of Covid-19 in males.

### Wrangling Data
```{r}
Males<-filter(FinalData,sex== "M") %>%
  select("treatment","infected")
```

Here I filtered the dataset so that only males, our subpopulation of interest, remain. I also selected only the columns needed to determine the effecicacy of the vaccine, resulting in a tidier dataset than I otherwise would have.

### Graphical Results
Since we are trying to explain a categorical variable, "infected", by another categorical variable, "treatment", we should use a boxplot to visualize the relationship between the two variables. we are also more interested in percentages than quantities.

```{r}
barchartGC(~infected + treatment,data=Males, type="percent")
```
Here we can see that the percentage of those who were vaccinated contracted covid-19 is significantly smaller than those who had taken placebo. Of those who did not contract covid-19, roughly half were vaccinated, while half were not.

### Numerical Results
We shall create a table showing the data, in a similar manner to the above plot.
```{r}
tableM <- xtabs(~infected + treatment,data=Males)
rowPerc(tableM)
```
Here we can see the same results as the bar plot, but more precisely. Of those who were infected with covid-19, 68.44% had taken placebo, while 31.56% had taken the vaccine. Of those who remained uninfected, close to half (49.88%) had taken placebo, while the other "half" (50.12%) had taken the vaccine.

### Inferential Results
Since we are explaining a categorical variable with another categorical variables, and we only are using two variables, we shall use the fisher test.
```{r}
fisher.test(tableM)
```
We have a p-value of 2.364e-12 (0.000000000002364), corresponding to a 0.0000000002364% likelihood that our results are due to random chance. This is highly improbable, thus we reject $H_0$ in favor of $H_A$

### Conclusion
We have rejected the null hypothesis, and thus concluded that the treatment does affect the probability of contracting covid-19 in males. The group of those who were uninfected is roughly half and half between vaccine and placebo, but this is because of a low probability of contracting the disease in general. Of those infected, 68.44% had taken placebo, while 31.56% had taken the vaccine. This means that, for males, those who take the vaccine are 2.17 times less likely to contract covid-19 than those who did not.


## Females

$H_0$ The vaccine covid921 is has no effect on the contraction of Covid-19 in females.

$H_1$ The vaccine covid921 is has an effect on the contraction of Covid-19 in females.

### Wrangling Data
```{r}
Females<-filter(FinalData,sex== "F") %>%
  select("treatment","infected")
```

Here I once again filtered the dataset so that only females, our subpopulation of interest, remain..


## LGBTQ



## Druggies


# Overall Results and Conclusions